---
title: "ANNs testing"
output: html_notebook
---

### Adjustment of artificial neural networks (ANNs) for the prediction of productivity of forwarders working in the Northeast of Corrientes
#### Author: Leszczuk Andrés

#### libraries
```{r warning=FALSE}
library(readxl)     #Read excel file
library(dplyr)      #Data analysis
library(neuralnet)  #fit neural net
library(relaimpo)   #Relative importance in Neural Net
library(sigmoid)    #For Relu function
library(ggplot2)
```
#### Data
```{r warning=FALSE}
# Import data
data <- data <- read_excel("path/0-data/bd_clean.xlsx")

# Select variables
data <- data[,c("Vol_i", "VOL_TOT", "PEF", "DE")]

# Filter out rows with NA values in any of the selected variables
data <- data %>% filter(!is.na(Vol_i) & 
                        !is.na(VOL_TOT) & 
                        !is.na(DE) &
                        !is.na(PEF))
```

#### Show data structure
```{r}
str(data)
```

#### Scaling of values to reduce variations between variables and improve the fitting performance of neural networks
```{r}
maxs <- apply(data, 2, max)
mins <- apply(data, 2, min)
scaled <- as.data.frame(scale(data, center = mins, scale = maxs - mins))
summary(scaled)
```
`#### Data sampling 
We separated the data into two random data sets, 70% for train (n = 295) and 30% for test data (n = 126 data).
```{r}
set.seed(1) # fit a seed 
indice <- sample(1:nrow(data), round(0.7*nrow(data)))


train <- data[indice,] # train data 
test <- data[-indice,] # test data

train_summary <- train %>% summarise(N = n(),
                    mean_DE = mean(DE, na.rm = T),
                    sd_DE   = sd(DE, na.rm = T),
                    min_DE = min(DE, na.rm = T),
                    max_DE = max(DE, na.rm = T),
                    
                    mean_Vol_i = mean(Vol_i, na.rm = T),
                    sd_Vol_i   = sd(Vol_i, na.rm = T),
                    min_Vol_i = min(Vol_i, na.rm = T),
                    max_Vol_i = max(Vol_i, na.rm = T),
                    
                    mean_VOL_TOT = mean(VOL_TOT, na.rm = T),
                    sd_VOL_TOT   = sd(VOL_TOT, na.rm = T),
                    min_VOL_TOT = min(VOL_TOT, na.rm = T),
                    max_VOL_TOT = max(VOL_TOT, na.rm = T))


test_summary <- test %>% summarise(N = n(),
                    mean_DE = mean(DE, na.rm = T),
                    sd_DE   = sd(DE, na.rm = T),
                    min_DE = min(DE, na.rm = T),
                    max_DE = max(DE, na.rm = T),
                    
                    mean_Vol_i = mean(Vol_i, na.rm = T),
                    sd_Vol_i   = sd(Vol_i, na.rm = T),
                    min_Vol_i = min(Vol_i, na.rm = T),
                    max_Vol_i = max(Vol_i, na.rm = T),
                    
                    mean_VOL_TOT = mean(VOL_TOT, na.rm = T),
                    sd_VOL_TOT   = sd(VOL_TOT, na.rm = T),
                    min_VOL_TOT = min(VOL_TOT, na.rm = T),
                    max_VOL_TOT = max(VOL_TOT, na.rm = T))

train_summary
test_summary


```


####` Scaled Data sampling 
We separated the data into two random data sets, 70% for train (n = 295) and 30% for test data (n = 126 data).
```{r}
set.seed(1) # fit a seed 
indice <- sample(1:nrow(data), round(0.7*nrow(data)))
test <- data[-indice,]

train_ <- scaled[indice,] # train data 
test_ <- scaled[-indice,] # test data
```
#### fitting of several artificial neural networks
After the tests used in the previous script, we fit the neural network that showed the best result in terms of coefficient of variation, RMSE and R² fit.

```{r}

#RELU <- function(x) relu(x)
#softplus <- function(x) log(1 + exp(x))

n <- neuralnet(PEF~Vol_i+VOL_TOT+DE,
                 data = train_,
                 threshold = 0.001,
                 act.fct = "logistic",
                 linear.output = T,
                 hidden = c(6,6,4),
                 stepmax = 1000000)
```
#### We plot the neural network diagram
```{r}
plot(n,rep = 1)
```
#### Test the neural net
we test the neural network from the test data 
With the test and train data 
```{r}
# with test data

pr.nn <- compute(n, test_) # This takes the test data and predicts the productivity of the forwarder.
#pr.nn$net.result           # The values are scaled
# are transformed to productivity values 
pr.nn_ <- pr.nn$net.result*(max(data$PEF)-min(data$PEF))+min(data$PEF)
test.r <- (test_$PEF)*(max(data$PEF)-min(data$PEF))+min(data$PEF)
MSE.nn <-sum((test.r - pr.nn_)^2)/nrow(test_)

summary(pr.nn_)

print(paste('RMSE',MSE.nn))
print(paste('cor',cor(test.r,pr.nn$net.result)))
```
With train data
```{r}
# with train

pr.nn.train <- compute(n, train_) # This takes the test data and predicts the productivity of the forwarder.
#pr.nn$net.result           # The values are scaled
# are transformed to productivity values 
pr.nn_.train <- pr.nn.train$net.result*(max(data$PEF)-min(data$PEF))+min(data$PEF)
test.r.train <- (train_$PEF)*(max(data$PEF)-min(data$PEF))+min(data$PEF)
MSE.nn_train <-sum((test.r.train - pr.nn_.train)^2)/nrow(train_)

summary(pr.nn_.train)

print(paste('RMSE',MSE.nn_train))
print(paste('cor',cor(test.r.train,pr.nn.train$net.result)))
```

#### Predicted vs. observed productivity plot
```{r, warning=FALSE}
db <- data.frame(pr.nn_, data[-indice,3])
db$dummy <- "dummy"
db$id <- 1:dim(db)[1]
colnames(db) <- c("fitted", "observed", "dummy","id")

summary(lm(observed~fitted, db))

library(ggplot2)

theme_crojfe <- function(base_size = 10, base_family = "Tahoma") {
  theme_classic(base_size = base_size, base_family = base_family) +
    theme(
      panel.border = element_rect(color = "#4D4D4D", fill = NA, linewidth = 0.4),
      axis.line = element_line(color = "#4D4D4D", linewidth = 0.4),
      axis.ticks = element_line(color = "#4D4D4D", linewidth = 0.4),
      
      # grid for CROJFE
      panel.grid.major = element_line(color = "#B3B3B3", linewidth = 0.35),
      panel.grid.minor = element_blank(),
      
      legend.position = "none",
      
      plot.title = element_text(hjust = 0.5, size = base_size),
      axis.title = element_text(size = base_size),
      axis.text  = element_text(size = base_size)
    )
}


# Gráfico de dispersión comparando valores reales y predichos
plot1_nn <- ggplot(db, aes(y = observed, x = fitted)) +
  geom_point(shape = 21, fill = "#CFCFCF", color = "#4D4D4D", size = 2, stroke = 0.3) +
  geom_smooth(method = "lm", formula = y ~ x, color = "#4D4D4D", linewidth = 0.6, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, color = "black", linewidth = 0.5) +
  labs(
    x = expression("Predicted productivity, m"^3*" h"^{-1}),
    y = expression("Observed productivity, m"^3*" h"^{-1})
  ) +
  scale_x_continuous(breaks = seq(0, 120, 20), limits = c(0, 120)) +
  scale_y_continuous(breaks = seq(0, 120, 20), limits = c(0, 120)) +
  #coord_equal() +
  theme_crojfe(base_size = 8, base_family = "Tahoma")


lm_eq <- lm(observed ~ fitted, data = db)

eq <- bquote(italic(y) == .(round(coef(lm_eq)[1], 2)) + .(round(coef(lm_eq)[2], 2)) %.% italic(x))
r_squared <- bquote(R^2 == .(round(summary(lm_eq)$r.squared, 2)))
sample_size <- bquote(italic(n) == .(nrow(test)))

plot_with_eq <- plot1_nn +
  annotate("text", x = 50, y = 25, label = as.expression(eq), family = "Tahoma", size = 2.5, hjust = 0) +
  annotate("text", x = 50, y = 17, label = as.expression(r_squared), family = "Tahoma", size = 2.5, hjust = 0) +
  annotate("text", x = 50, y = 9,  label = as.expression(sample_size), family = "Tahoma", size = 2.5, hjust = 0)


# Mostrar el gráfico con la ecuación, R² y el valor de 'n'
print(plot_with_eq)
```
# Residual plot 
```{r}
residuos <- db$observed - db$fitted

plot2 <- ggplot(data.frame(residuos), aes(x = residuos)) +
  geom_density(fill = "#CFCFCF", color = "black", linewidth = 0.5) +
  labs(
    x = "Residuals",
    y = "Density"
  ) +
  theme_crojfe(base_size = 8, base_family = "Tahoma")
residuos_df <- data.frame(PEF_observed = test$PEF, Residuos = residuos)

```

Grafico de residuos
```{r}

residuos_df <- data.frame(PEF_observed = test$PEF, Residuos = residuos)

plot3 <- ggplot(residuos_df, aes(x = PEF_observed, y = Residuos)) +
  geom_point(shape = 21, fill = "#CFCFCF", color = "#4D4D4D", size = 2, stroke = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.5) +
  labs(
    x = expression("Observed Productivity, m"^3*" h"^{-1}),
    y = "Residuals"
  ) +
  theme_crojfe(base_size = 8, base_family = "Tahoma")

```

```{r, warning=FALSE}
library(ggpubr)
PLOT_NN <- ggarrange(
  plot_with_eq, plot2, plot3,
  labels = c("A", "B", "C"),
  font.label = list(size = 8, face = "bold", family = "Tahoma"),
  ncol = 3,
  widths = c(1, 1, 1),   # <- fuerza mismo ancho
  align = "hv"
)
PLOT_NN
ggsave("Fig4.png",PLOT_NN, width = 18, height = 9, units = "cm", dpi = 600)
```


# Grafico con todos los datos para predichos vs observados e histograma 
```{r}
scaled <- scaled[,-5]
scaled

prueba.nn <- compute(n, scaled)
prueba.nn.pef <- prueba.nn$net.result*(max(data$PEF)-min(data$PEF))+min(data$PEF)

db_ <- data.frame(prueba.nn.pef, data[,3])
db_$dummy <- "dummy"
db_$id <- 1:dim(db_)[1]
colnames(db_) <- c("fitted", "observed", "dummy","id")

# filer outliers

#db_ <- filter(db_, id !=115)
#db_ <- filter(db_, id !=95)
#db_ <- filter(db_, id !=253)
#db_ <- filter(db_, id !=116)
#db_ <- filter(db_, id !=244)
#db_ <- filter(db_, id !=254)
#db_ <- filter(db_, id !=270)
#db_ <- filter(db_, id !=6)
#db_ <- filter(db_, id !=416)
#db_ <- filter(db_, id !=407)
#db_ <- filter(db_, id !=388)
#db_ <- filter(db_, id !=240)
#db_ <- filter(db_, id !=200)
#db_ <- filter(db_, id !=357)

summary(lm(fitted~observed, db_))

p6 <- ggplot(db_ , aes(fitted, observed,  color = dummy, label = id))+
  geom_point(shape = 21, fill = "lightgray", size = 3.5)+
  geom_smooth(method = "lm", formula = y~x, se = F)+
  geom_abline(intercept = 0, slope = 1)+
  #geom_text()+
  theme(panel.grid.major = element_line(colour = "black"),
        panel.border = element_rect(linetype = 1, fill = NA),
        panel.background = element_rect(fill = "white"),
        legend.position = c(.95, .25),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        text = element_text(size = 12))+
  scale_fill_manual(values = c('#757474'))+
  scale_colour_manual(values = c('#757474'))+
  annotate(geom = "text", x = 35, y = 110, label = "R² = 0.84")+
  expand_limits(x=c(10,50),y=c(10,50))+
  scale_x_continuous(breaks = seq(0,120,10))+
  scale_y_continuous(breaks = seq(0,120,10))+
  xlab("Effective productivity fitted , m³/h")+
  ylab("Effective productivity observed, m³/h")+
  theme(legend.position = "none")
p6
```
#### Residual plot
```{r}

mod <- lm(observed~fitted, db_)
residuals(mod)
standard_res <- rstandard(mod)

db_$residuals <- standard_res

p4 <- ggplot(db_, aes(residuals))+
  xlab("Residuals, m³/h")+
  ylab("Frecuency of residuals")+
  theme(panel.grid.major = element_line(colour = "black"),
        panel.border = element_rect(linetype = 1, fill = NA),
        panel.background = element_rect(fill = "white"),
        legend.position = c(.95, .25),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6),
        text = element_text(size = 12))+
  scale_fill_manual(values = c('#757474'))+
  scale_colour_manual(values = c('#757474'))+
  theme(legend.position = "none")+
  geom_histogram(bins = 25)
  
p4
```
```{r}
# unificate 2 plots in one
library(ggpubr)
p64 <- ggarrange(p6,p4,
                 labels = c("C", "D"),
                 ncol = 2)
p64
```